---
title: "Project_Spark_Fund"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Scope of work**

Spark Funds is an asset management company. Spark Funds wants to make investments in a few companies. The CEO of Spark Funds wants to understand the global trends in investments so that she can take the investment decisions effectively.

Spark Funds has two constraints for investments:
* It wants to invest between 5 to 15 million USD per round of investment.
* It wants to invest only in English-speaking countries because of the ease of communication with the companies it would invest in.

**Goals of data analysis**

* Investment type analysis: Understanding investments in venture, seed/angel, private equity categories, etc. so Spark Funds can decide which type is best suited for its strategy.
* Country analysis: Understanding which countries have had the most investments in the past. These will be Spark Funds' favourites as well.
* Sector analysis: Understanding the distribution of investments across the eight main sectors. (Note that we are interested in the eight main sectors provided in the mapping file. The two files - companies and rounds2 - have numerous sub-sector names; hence, you will need to map each sub-sector to its main sector.)

**Data Dictionary**

**companies: A table with basic data of companies**

* Permalink - Unique ID of company
* name - Company name
* homepage_url - Website URL
* category_list - Category/categories to which a company belongs
* status - Operational status
* country_code - Country Code 
* state_code - State

**rounds2: The most important parameters are explained below**

* company_permalink - Unique ID of company
* funding_round_permalink - Unique ID of funding round
* funding_round_type - Type of funding (venture, angel, private equity etc)
* funding_round_code - Round of venture funding (round A, B etc.)
* funded_at - Date of funding
* raised_amount_usd - Money raised in funding (USD)

**Questions**

* How many unique companies are present in rounds2?
* How many unique companies are present in companies?
* Merge the two data frames so that all variables (columns) in the companies frame are added to the rounds2 data frame. Name the merged frame master_frame. How many observations are present in master_frame?
* How many NA values are present in the column raised_amount_usd?
* What do you replace NA values of raised_amount_usd with? Enter a numeric value.
* Average funding amount of venture type?                             
* Average funding amount of angel type?
* Average funding amount of seed type?
* Average funding amount of private equity type?
* Considering that Spark Funds wants to invest between 5 to 15 million USD per investment round, which investment type is the most suitable for it?
* Total number of investments (count)?
* Total amount of investment (USD)?
* Top sector (based on count of investments)?
* Second-best sector (based on count of investments)?
* Third-best sector (based on count of investments)?
* Top two companies to which investements were rendered for the top three countries?


Setting up the working directory and loading the required files
```{r}
companies <- read.delim("companies.txt",header = TRUE, sep = "\t",stringsAsFactors = FALSE)
round2 <- read.csv("rounds2.csv",header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

Checking the structure of companies and round2 dataframes.
```{r}
str(companies)
str(round2)
```

Loading the required library
```{r warning=FALSE , message=FALSE}
library(dplyr)
library(tidyr)
```

As company_permalink column in round2 dataframe has mixed cases for the same permalink, hence converting conpamy_permalink column to uppercase, storing it in upper_permalink and then getting the unique values.

We will use the mutate function from dplyr package to create the upper_permalink variable for round2 dataframe.
```{r}
round2_mutate <- mutate(round2, upper_permalink = toupper(round2$company_permalink))
dim(round2_mutate)
```

How many unique companies are present in rounds2?

Using length function to find the unique permalink.
```{r}
unique_round2 <- (length(unique(round2_mutate$upper_permalink)))
unique_round2
```

How many unique companies are present in companies?

Using permalink from companies dataframe to find unique companies
```{r}
unique_companies <- (length(unique(companies$permalink)))
unique_companies
```

Before merging companies and round2 dataframe to make master dataframe we have converted company_permalink and permalink to all caps so that the merging can be done properly
```{r}
companies[,1] = toupper(companies[,1])
round2[,1] = toupper(round2[,1])
```

Merge the two data frames so that all variables (columns) in the companies frame are added to the rounds2 data frame. Name the merged frame master_frame. How many observations are present in master_frame?

Creating the master dataframe.
We will be using the left join function from dplyr package to do the merging
```{r}
master_frame <- left_join(round2,companies,by = c("company_permalink" = "permalink"))
dim(master_frame)
```

How many NA values are present in the column raised_amount_usd?

Finding NA in raised amount usd column of master_frame
```{r}
number_of_na <- sum(is.na(master_frame$raised_amount_usd))
number_of_na
```

What do you replace NA values of raised_amount_usd with? Enter a numeric value.

Replacing NA with zero

**Please Note:- It is not advisable to replace NA with zero, it is always better to keep the NA's as is.**
```{r}
master_frame$raised_amount_usd[is.na(master_frame$raised_amount_usd)] <- 0
```

Finding the average of funding type.

Average funding amount of venture type?                             
Average funding amount of angel type?
Average funding amount of seed type?
Average funding amount of private equity type?

This will tell us about the average funding amount for venture, angel, seed & private equity

We will be using the group_by and summarise functions to find the average of funding types
```{r}
average_of_funding_types <- summarise(group_by(master_frame, funding_round_type), mean(raised_amount_usd,na.rm = TRUE))
colnames(average_of_funding_types)[2] <- "Average Amount"
average_of_funding_types
```

Considering that Spark Funds wants to invest between 5 to 15 million USD per investment round, which investment type is the most suitable for it?

Converting to millions to find what funding type falls within 5 to 15 million. This conversion will easily show what type falls into the 5 to 15 million bucket.
```{r}
convert_to_millions <- average_of_funding_types$`Average Amount`/1000000
average_of_funding_types_in_millions <- cbind(average_of_funding_types, convert_to_millions)
average_of_funding_types_in_millions
```

Looking at the above table we can see that venture type is the one where spark should invest.Hence subsetting master_frame to get line items related to only venture fund type

```{r}
venture_fund_type <- subset(master_frame, funding_round_type == "venture")
```

Grouping and summarizing the venture fund type dataframe on a counrty code level by the basis of fund raised
```{r}
fundraised_by_countrycode <- summarise(group_by(venture_fund_type, country_code), sum(raised_amount_usd,na.rm = TRUE))
colnames(fundraised_by_countrycode)[2] <- "Total Amount"
fundraised_by_countrycode
```

After we have summarized venture fund type on a country_code level its time to get the top 9 by fund raised, this is just the raw table

We will be using the top_n function from dplyr package for this insight.
```{r}
top_9_raw <- top_n(fundraised_by_countrycode,9)
```

Converting Total Amount to millions and billions for better readability
```{r}
convertmillion <- top_9_raw$`Total Amount`/1000000
convertbillion <- top_9_raw$`Total Amount`/1000000000
```

Here we are attaching convertmillion & convertbillion to top_9_raw to form a new dataframe called top_9
```{r}
top_9 <- cbind(top_9_raw,convertmillion,convertbillion)
arrange(top_9,desc(convertbillion))
```

Now we will load the sector mapping file to R console
```{r}
setwd("C:/Users/mishrsat/Desktop/Automation/codes/Project 1 Spark Fund")
sector_mapping_raw <- read.csv("mapping.csv",header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

Using the gather function we are converting the wide format data to long format data
```{r}
sector_mapping_gather <- gather(sector_mapping_raw, Catagory, Catagory_value, Automotive...Sports:Social..Finance..Analytics..Advertising)
```

Removing all the row that has a value of 0 in the newly created catagory_value column as row that have 1 is what we require
```{r}
sector_mapping_gather <- sector_mapping_gather[!(sector_mapping_gather$Catagory_value == 0),]
```

Removing the third column (Catagory_value) from the dataframe as it dosen't add any value any more
```{r}
sector_mapping_gather <- sector_mapping_gather[, -3]
```

Using the gsub function, here we are replacing the characters which have typo errors in the mapping master. There are a total of four character strings that were replaced
```{r}
sector_mapping_gather$category_list <- gsub("0","na",sector_mapping_gather$category_list)
sector_mapping_gather$category_list <- gsub("nanotechnology","Nanotechnology",sector_mapping_gather$category_list)
sector_mapping_gather$category_list <- gsub("natural","Natural",sector_mapping_gather$category_list)
sector_mapping_gather$category_list <- gsub("navigation","Navigation",sector_mapping_gather$category_list)
```

Mutating the column category_list to remove multipe category in each row. This is done as per the business rule set up by spark fund
```{r}
venture_fund_type <- mutate(venture_fund_type,category_list=sapply(strsplit(venture_fund_type$category_list, split='|', fixed=TRUE),function(x) (x[1])))
```

Joining venture_fund_type and sector_mapping_gather to create the final cleansed dataframe

Using the left_join function from dplyr package to make the merge
```{r}
Venture_fund_final <- left_join(venture_fund_type,sector_mapping_gather,by = c("category_list" = "category_list"))
```

* Using the venture fund final dataframe we are creating a new vector raised_usd_in_millions to store values in millions
* Then we are combining the vector to venture fund final to form combined raw dataframe
* After that we are taking the rows that fall in the bucket of 5 to 15 million and creating combined_final
```{r}
raised_usd_in_millions <- Venture_fund_final$raised_amount_usd/1000000
combined_raw <- cbind(Venture_fund_final,raised_usd_in_millions)
combined_final <- combined_raw[(combined_raw$raised_usd_in_millions >= 5.0) & (combined_raw$raised_usd_in_millions <= 15.0) & !is.na(combined_raw$raised_usd_in_millions), ]
```

Now we are subsetting and creating three data frames , one each for the top three countries
```{r}
united_states <- subset(combined_final, country_code == "USA")
great_britain <- subset(combined_final, country_code == "GBR")
india <- subset(combined_final,country_code == "IND")
```

Total number of investments (count)?

Using the tally function we will first get the count and then sort it
```{r}
total_countofinvestment <- tally(group_by(combined_final,country_code), sort = TRUE)
head(total_countofinvestment,10)
```

Total amount of investment (USD)? 

Using the summarise and group_by function we are getting the amount on a counrty code level
```{r}
total_amtofinvestment <- summarise(group_by(combined_final, country_code),sum(raised_amount_usd))
```

Top sector (based on count of investments)?
Second-best sector (based on count of investments)?
Third-best sector (based on count of investments)?

Finding the top 3 sectors and count of respective sector. We will be using the tally function to get the count and then sort it
```{r}
sector_name_count_USA <- tally(group_by(united_states,Catagory), sort = TRUE)
sector_name_count_GRB <- tally(group_by(great_britain,Catagory), sort = TRUE)
sector_name_count_IND <- tally(group_by(india,Catagory), sort = TRUE)
sector_name_count_USA
sector_name_count_GRB
sector_name_count_IND
```

**Finding the top two companies to which investements were rendered for USA, GBR & IND**

* Here for the top company in USA we first summarise and group_by Catagory & name on united_states dataframe
* From there we subset the dataframe on others catagory
* Then we again summarise and group_by on name from the subsetted dataframe to get the total amount given to each company
* Using the top_n function we find the top most company
```{r}
first_high_invest_USA <- summarise(group_by(united_states, Catagory,name), sum(raised_amount_usd))
USA_others <- subset(first_high_invest_USA,Catagory == "Others")
USA_others_summary <- summarise(group_by(USA_others,name), sum(`sum(raised_amount_usd)`))
first_high_invest_USA_final <- top_n(USA_others_summary,1)
first_high_invest_USA_final
```

* Here for the top company in USA we first summarise and group_by Catagory & name on united_states dataframe
* From there we subset the dataframe on Social_Finance_Analytics_Advertising catagory
* Then we again summarise and group_by on name from the subsetted dataframe to get the total amount given to each company
* Using the top_n function we find the top most company
```{r}
second_high_invest_USA <- summarise(group_by(united_states, Catagory,name), sum(raised_amount_usd))
USA_SFAA <- subset(second_high_invest_USA,Catagory == "Social..Finance..Analytics..Advertising")
USA_SFAA_summary <- summarise(group_by(USA_SFAA,name), sum(`sum(raised_amount_usd)`))
second_high_invest_USA_final <- top_n(USA_SFAA_summary,1)
second_high_invest_USA_final
```

* Here for the top company in GBR we first summarise and group_by Catagory & name on great_britain dataframe
* From there we subset the dataframe on others catagory
* Then we again summarise and group_by on name from the subsetted dataframe to get the total amount given to each company
* Using the top_n function we find the top most company
```{r}
first_high_invest_GRB <- summarise(group_by(great_britain, Catagory,name), sum(raised_amount_usd))
GRB_others <- subset(first_high_invest_GRB,Catagory == "Others")
GRB_others_summary <- summarise(group_by(GRB_others,name), sum(`sum(raised_amount_usd)`))
first_high_invest_GRB_final <- top_n(GRB_others_summary,1)
first_high_invest_GRB_final
```

* Here for the top company in GBR we first summarise and group_by Catagory & name on great_britain dataframe
* From there we subset the dataframe on Social_Finance_Analytics_Advertising catagory
* Then we again summarise and group_by on name from the subsetted dataframe to get the total amount given to each company
* Using the top_n function we find the top most company
```{r}
second_high_invest_GRB <- summarise(group_by(great_britain, Catagory,name), sum(raised_amount_usd))
GRB_SFAA <- subset(second_high_invest_GRB,Catagory == "Social..Finance..Analytics..Advertising")
GRB_SFAA_summary <- summarise(group_by(GRB_SFAA,name), sum(`sum(raised_amount_usd)`))
second_high_invest_GRB_final <- top_n(GRB_SFAA_summary,1)
second_high_invest_GRB_final
```

* Here for the top company in IND we first summarise and group_by Catagory & name on india dataframe
* From there we subset the dataframe on others catagory
* Then we again summarise and group_by on name from the subsetted dataframe to get the total amount given to each company
* Using the top_n function we find the top most company
```{r}
first_high_invest_IND <- summarise(group_by(india, Catagory,name), sum(raised_amount_usd))
IND_others <- subset(first_high_invest_IND,Catagory == "Others")
IND_others_summary <- summarise(group_by(IND_others,name), sum(`sum(raised_amount_usd)`))
first_high_invest_IND_final <- top_n(IND_others_summary,1)
first_high_invest_IND_final
```

* Here for the top company in IND we first summarise and group_by Catagory & name on india dataframe
* From there we subset the dataframe on Social_Finance_Analytics_Advertising catagory
* Then we again summarise and group_by on name from the subsetted dataframe to get the total amount given to each company
* Using the top_n function we find the top most company
```{r}
second_high_invest_IND <- summarise(group_by(india, Catagory,name), sum(raised_amount_usd))
IND_SFAA <- subset(second_high_invest_IND,Catagory == "Social..Finance..Analytics..Advertising")
IND_SFAA_summary <- summarise(group_by(IND_SFAA,name), sum(`sum(raised_amount_usd)`))
second_high_invest_IND_final <- top_n(IND_SFAA_summary,1)
second_high_invest_IND_final
```

**THIS PROJECT MAINLY DEALT WITH SLICING AND DICING OF DATA USING R LANGUAGE**

















